---
title: "Import Coastal Water Quality Data"
author: "Francois van Schalkwyk"
date: "2025-03-19"
output: html_document
---

# Workspace Set-Up

Make sure that you are working in the outfalls project, found in the working directory.

## Set working directory for all code chunks to project root directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Load required packages

```{r message = FALSE}
library(tidyverse)
library(dbplyr)
library(DBI)
library(pdftools)
library(readxl)
```

## Database Connection

```{r db connection}
# Enter credentials when prompted
con <- dbConnect(RPostgres::Postgres(),
  host = "aws-0-eu-central-1.pooler.supabase.com",
  port = 5432,
  user = str_c(rstudioapi::askForPassword(prompt = "Username: "), "wnrvdvesovnkmqbhkhjz", sep = "."),
  password = rstudioapi::askForPassword(),
  dbname = "postgres"
)
```

```{r}
dbGetInfo(con)
```

# CMB Spreadsheets

```{r list-files}
# List the files in the directory
path <- "01 Input Data/Coastal Water Quality Data/"
files <- list.files(path = path, pattern = "*.xls*")
full_path <- paste(path, files, sep = "")
```

```{r read-excel-function-copy}
# Create function for extracting data and populating table
read_bacteria_data <- function(file, sheet){
  # Extract filename
  filename <- file |> str_remove("01 Input Data/Coastal Water Quality Data/")
  
  # Read Ecoli data
  data <- read_excel(file,
                      range = "A1:B1000",
                      sheet = sheet,
                      col_names = c("date", "value_cfu_per_100ml"),
                      col_types = c("date", "numeric"))
  
  # If empty tibble, create new empty tibble
  if (nrow(data) == 0) {
    data <- tibble(date = date(), value_cfu_per_100ml = numeric())
    data <- add_case(data, date = NA, value_cfu_per_100ml = NA) |>
      mutate(date = date(date))
  }
  
  results <- data |>
    slice(-1) |>
    filter(!is.na(value_cfu_per_100ml)) |>
    mutate(
      date = date(date),
      file = filename
      )

  return(results)
}
```

```{r run-function}
# Run the function for faecal streptococci
faecal <- str_c(path, files) |>
  map_df(~read_bacteria_data(.x, 3))
```

```{r summary-statistics}
# Check summary statistics for faecal streptococci
faecal |>
  group_by(file) |>
  arrange(file) |>
  summarise(
    min_date = min(date),
    max_date = max(date),
    n = sum(!is.na(value_cfu_per_100ml)),
    min = min(value_cfu_per_100ml, na.rm = TRUE),
    max = max(value_cfu_per_100ml, na.rm = TRUE)
    )
```

*Faecal streptococci:*
Some clear threshold values: 2419, 7257, 15000, 4500, 24196, 450, 72588
Some clear divided data: 0.033, 0.33 and then some strange decimal data but very consistent: 0.1249387

```{r faecal-plot}
faecal |>
  ggplot(aes(x = date, y = value_cfu_per_100ml, colour = file)) +
  geom_point() +
  theme(legend.position = "none") +
  coord_cartesian(xlim = c(ymd("2014-01-01"), ymd("2025-06-01")))
```

The faecal streptococci data don't seem to have the same solid lines that the E. coli data had. There are also some streaks but these streaks occur at frequencies that can be reasonably expected from weekly sampling. These are also likely to be due to censored values, which will be dealt with in due course.

The values greater than 70000 come from Enterolert values that were diluted and still exceeded the threshold. The threshold value of 2419 was exceeded, the value was diluted 10 times, and it still exceeded the 24190 value. This was multiplied by 3 to get 72000.

I want to separate out the data that are in one sheet but not in another. 
Let me start with one file and see how it goes.

```{r added-data}
faecal <- faecal |>
  mutate(
    site_id = str_extract(file, "[:upper:]{2,3}\\d{2}[:upper:]?"),
    added = str_detect(file, "SSB|CMB")
      ) |>
  select(-file)

# Which records have been added to the spreadsheets and are not in the SSB spreadsheets?
added <- dplyr::setdiff(
  faecal |> filter(added) |> select(-added),
  faecal |> filter(!added) |> select(-added)
  ) |>
  arrange(date, site_id)

# Combine all the data and remove the duplicates
all <- dplyr::union(
  faecal |> filter(added) |> select(-added),
  faecal |> filter(!added) |> select(-added)
)

# Keep only the SSB data
only_ssb <- dplyr::setdiff(all, added)

# Combine the data sets again
faecal <- bind_rows(
  added |> mutate(source = "CMB"),
  only_ssb |> mutate(source = "SSB")
)
```

```{r censored-values}
# Replace censored values with character value including < operator
faecal <- faecal |>
  mutate(censored_value = case_when(
    round(value_cfu_per_100ml, 2) == 0.67 ~ "<2",
    (round(value_cfu_per_100ml, 2) == 0.33 | value_cfu_per_100ml == 0.3) ~ "<1",
    round(value_cfu_per_100ml, 3) == 0.033 ~ "<1",
    round(value_cfu_per_100ml, 2) == 0.12 ~ "<1",
    round(value_cfu_per_100ml, 2) == 0.22 ~ "<1",
    round(value_cfu_per_100ml, 2) == 3.33 ~ "<10",
    round(value_cfu_per_100ml, 2) == 16.67 ~ "<50",
    round(value_cfu_per_100ml, 2) == 33.33 ~ "<100",
    round(value_cfu_per_100ml, 2) == 806.67 ~ ">2420",
    round(value_cfu_per_100ml, 2) == 3.48 ~ "4",
    round(value_cfu_per_100ml, 2) == 4.18 ~ "4",
    round(value_cfu_per_100ml, 2) == 2.48 ~ "2",
    value_cfu_per_100ml == 450 ~ ">150",
    value_cfu_per_100ml == 4500 ~ ">1500",
    value_cfu_per_100ml == 15000 ~ ">5000",
    value_cfu_per_100ml == 7257 ~ ">2419",
    value_cfu_per_100ml == 1.3 ~ "ND",
    value_cfu_per_100ml == 72588 ~ ">24196",
    value_cfu_per_100ml == 24196 ~ ">24196",
    value_cfu_per_100ml == 15000 ~ ">5000",
    value_cfu_per_100ml == 1500 & date > "2023-08-01" ~ ">1500",
    value_cfu_per_100ml == 7257 ~ ">2419",
    value_cfu_per_100ml == 450 & date > "2023-08-01" & date < "2024-11-01" ~ ">150",
    .default = as.character(value_cfu_per_100ml)
    )) |>
  mutate(censored_value = if_else(censored_value == "0", "ND", censored_value))
```

Censored values can be found both at the lower detection limits and at the upper detection limits. Streaks at the upper limits can also indicate censored values

```{r upper-censored-values-faecal}
faecal |>
  ggplot(aes(x = date, y = value_cfu_per_100ml)) +
  geom_point(alpha = 0.5) +
  theme(
    legend.position = "none"
  ) +
  coord_cartesian(xlim = c(ymd("2014-01-01"), ymd("2024-11-01"))) +
  scale_x_date(minor_breaks = "1 year") +
  facet_grid(cols = vars(source))
```

The obvious streaks that were divisible by 3 were divided by three and preceded by the > operator. Other streaks were visible in the 2023 period that are not a consistent value and cannot be easily divided by 3, and are not therefore obvious censored values. These values are of concern, as the frequency indicates a non-random distribution.
Values are quite commonly distributed along 50 unit intervals. From mid-2020 to end of 2021 the value of 600 was very common and was replace during this period with <200 because this is common censored value.

For faecal streptococci, similar censored values were replaced with their character equivalents.

```{r check-censored}
faecal |>
  distinct(censored_value) |>
  arrange(censored_value)
```

```{r plot-censored}
faecal |>
  mutate(censored = str_detect(censored_value, "<|>")) |>
  ggplot(aes(x = date, y = value_cfu_per_100ml, colour = censored)) +
    geom_point() +
  coord_cartesian(xlim = c(ymd("2014-01-01"), ymd("2025-01-01")), ylim = c(0, 80000)) +
  scale_x_date(minor_breaks = "1 year")
```

```{r summary-statistics}
# Check summary statistics for faecal streptococci
faecal |>
  group_by(site_id) |>
  arrange(site_id) |>
  summarise(
    min_date = min(date),
    max_date = max(date),
    n = sum(!is.na(value_cfu_per_100ml)),
    min = min(value_cfu_per_100ml, na.rm = TRUE) |> as.integer(),
    max = max(value_cfu_per_100ml, na.rm = TRUE) |> as.integer()
    )

# Plot by site
faecal |>
  mutate(month = floor_date(date, unit = "months")) |>
  ggplot(aes(x = month, y = value_cfu_per_100ml)) +
  geom_col() +
  scale_x_date(minor_breaks = "1 year")
```

```{r save-data}
faecal |>
  mutate(numeric_value = case_when(
    censored_value |> str_detect("<") ~ censored_value |> str_remove("<|\\s") |> as.numeric() / 3,
    censored_value |> str_detect(">") ~ censored_value |> str_remove(">|\\s") |> as.numeric() * 3,
    censored_value |> str_detect("ND") ~ 0,
    .default = censored_value |> as.numeric()
  )) |>
  write_csv("02 Output Data/01 CSV/faecal_streptococci_data_clean.csv")
```

# SABS Data

In 2023/2024 data samples were collected with sample codes that don't match any other sample codes in the monitoring programme. I want to see how well our data corresponds to City of Cape Town data.

```{r read-sabs}
# Read in SABS data from Infinity
sabs <- read_excel("../../Infinity/Infinity Projects - Files/190-56 Coastal WQ/07_Cleaned Data/Coastal WQ Weekly 2023-24_cleaned.xlsx",
           col_names = c(
             "site_description",
             "cct_code",
             "cct_code_updated",
             "updated",
             "coords",
             "area",
             "week",
             "sample_date",
             "temp_c",
             "sabs_code",
             "enterococci_cfu_per_100ml",
             "enterococci_cfu_per_100ml_numeric",
             "sabs_sheet",
             "issues_flagged",
             "comments"
           ), col_types = c(
             "text",
             "text",
             "text",
             "text",
             "text",
             "text",
             "numeric",
             "date",
             "numeric",
             "text",
             "text",
             "numeric",
             "text",
             "text",
             "text"
           ), 
           skip = 1)

# Unique sites and descriptions
walab <- tbl(con, I("coastal.sites")) |> collect()
sabs_sites <- sabs |>
  distinct(cct_code, cct_code_updated, site_description) |>
  arrange(cct_code)

walab |>
  full_join(sabs_sites, by = c("site_id" = "cct_code_updated")) |>
  select(site_id, site_description.x, site_description.y)

# Summarise data
sabs |>
  summary()

# Check when coastal data end
tbl(con, I("coastal.results")) |>
  arrange(sample_date) |>
  arrange(sample_date)

# Export clean data to file
sabs |>
  select(cct_code_updated, sample_date, enterococci_cfu_per_100ml, enterococci_cfu_per_100ml_numeric, sabs_sheet) |>
  mutate(
    cct_code_updated = fct(cct_code_updated),
    sabs_sheet = str_c(sabs_sheet, ".pdf", sep = "")
    ) |>
  filter(!is.na(sample_date)) |>
  rename(site_id = cct_code_updated, censored_value = enterococci_cfu_per_100ml, filename = sabs_sheet) |>
  mutate(numeric_value = case_when(
    censored_value |> str_detect("<") ~ censored_value |> str_remove("<|\\s") |> as.numeric() / 3,
    censored_value |> str_detect(">") ~ censored_value |> str_remove(">|\\s") |> as.numeric() * 3,
    censored_value |> str_detect("ND") ~ 0,
    .default = censored_value |> as.numeric()
  )) |>
  select(sample_date, site_id, censored_value, filename) |>
  write_csv("02 Output Data/01 CSV/sabs_data_clean.csv")

# Compare SABS sites to WALAB sites
sabs
```

```{r join-data}
# Create simple table for SABS data that aligns with CCT data
sabs2023 <- sabs |>
  select(cct_code_updated, sample_date, enterococci_cfu_per_100ml, enterococci_cfu_per_100ml_numeric, week) |>
  mutate(cct_code_updated = fct(cct_code_updated)) |>
  filter(!is.na(sample_date)) |>
  rename(site_id = cct_code_updated, censored_value = enterococci_cfu_per_100ml) |>
  mutate(numeric_value = case_when(
    censored_value |> str_detect("<") ~ censored_value |> str_remove("<|\\s") |> as.numeric() / 3,
    censored_value |> str_detect(">") ~ censored_value |> str_remove(">|\\s") |> as.numeric() * 3,
    censored_value |> str_detect("ND") ~ 0,
    .default = censored_value |> as.numeric()
  )) |>
  select(sample_date, site_id, censored_value, numeric_value)

# How many records for each site?
filter(sabs2023) |>
  group_by(site_id) |>
  count() |>
  arrange(-n)

# Create a table with data for the same period as the SABS data
faec2023_full <- faecal |>
  filter(date >= "2023-10-03" & date <= "2024-11-06" & full) |>
  rename(sample_date = date, enterococci_cfu_per_100ml = censored_value) |>
  select(site_id, sample_date, enterococci_cfu_per_100ml) |>
  mutate(enterococci_cfu_per_100ml = if_else(enterococci_cfu_per_100ml == "<1", "ND", enterococci_cfu_per_100ml))

faec2023_ssb <- faecal |>
  filter(date >= "2023-10-03" & date <= "2024-11-06" & !full) |>
  rename(sample_date = date, enterococci_cfu_per_100ml = censored_value) |>
  select(site_id, sample_date, enterococci_cfu_per_100ml) |>
  mutate(enterococci_cfu_per_100ml = if_else(enterococci_cfu_per_100ml == "<1", "ND", enterococci_cfu_per_100ml))

# How many sites for each?
sabs2023 |>
  distinct(site_id) # 36 sites

faec2023_full |>
  distinct(site_id) # 20 sites

faec2023_ssb |>
  distinct(site_id) # 62 sites

# Find sites that do not have a combined spreadsheet
sites_not_combined <- setdiff(distinct(faec2023_ssb, site_id), distinct(faec2023_full, site_id)) |> pull(1)

# Number of records
nrow(sabs2023) # 1549
nrow(faec2023) # 1877

# Combine the three data sets
faec2023_full |> mutate(dataset = "SSB + CMB") |>
  bind_rows(sabs2023 |> mutate(dataset = "SABS")) |>
  bind_rows(faec2023_ssb |> mutate(dataset = "SSB")) |>
  arrange(sample_date, site_id, dataset, enterococci_cfu_per_100ml) |>
  filter(!(site_id %in% sites_not_combined & dataset == "SSB"))

sabs |>
  distinct(cct_code, cct_code_updated, site_description) |>
  arrange(cct_code)

# Handle censored values the same way
sabs <- sabs |>
  mutate(numeric = case_when(
    str_detect(enterococci_cfu_per_100ml, "<") ~ 0,
    str_detect(enterococci_cfu_per_100ml, ">") ~ as.numeric(str_remove(enterococci_cfu_per_100ml, ">")) * 3,
    str_detect(enterococci_cfu_per_100ml, "ND") ~ 0,
    .default = as.numeric(enterococci_cfu_per_100ml)
  ),
  sample_date = date(sample_date))

faec2023 <- faec2023 |>
  mutate(numeric = case_when(
    str_detect(enterococci_cfu_per_100ml, "<") ~ 0,
    str_detect(enterococci_cfu_per_100ml, ">") ~ as.numeric(str_remove(enterococci_cfu_per_100ml, ">")) * 3,
    str_detect(enterococci_cfu_per_100ml, "ND") ~ 0,
    .default = as.numeric(enterococci_cfu_per_100ml)
  ))

ggplot() +
  geom_point(data = sabs, aes(x = sample_date, y = numeric, colour = "SABS Data")) +
  geom_point(data = faec2023, aes(x = sample_date, y = numeric, colour = "CCT Data")) +
  theme(
    legend.position = "none"
  )

sabs |>
  full_join(faec2023, by = c("sample_date" = "date", "cct_code_updated" = "site_id"), keep = TRUE) |>
  select(cct_code_updated, sample_date, site_id, date, enterococci_cfu_per_100ml, value_cfu_per_100ml) |>
  filter(!if_any(c(cct_code_updated, site_id), is.na))

faecal |>
  filter(value_cfu_per_100ml > 60000)

filter(faecal, date == "2024-10-09" & site_id == "CN22")

```

Possible scenarios:
1. Exists in CCT SSB + CMB data + not in CCT SSB data + not in SABS data (SSB data that is not the SABS data)(sample date wrong in Excel)
2. Exists in CCT SSB + CMB data + not in CCT SSB data + exists in SABS data (SSB data that is from SABS data)(ideal)
3. No SSB + CMB sheet + not in CCT SSB + exists in SABS (No sheet for weekly samples)
4. No sheet in CCT + exists in SABS (site code not part of larger monitoring programme, doesn't correspond to current site code)
5. Not in CCT SSB + CMB + not in CCT SSB data + exists in SABS data (not captured)(not ideal)
6. Exists in CCT SSB + CMB + exists in CCT SSB + not in SABS data (SSB Data)(ideal)

The data set from Infinity has 1565 records, and for the same period, the CCT data set has 1877 records.
There are 569 combinations of site and date that exist in both data sets. There is some agreement between the values - many of the values of 3.33333 or 0 are ND values in the Infinity data. Some values do not agree between 
SABS data has 1565 records



# LIMS Data

```{r read-lims-data}
lims <- read_excel("01 Input Data/Coastal Water Quality Data/LIMS Data/LIMS Data.xlsm",
           col_names = c("sample_date", "site_id", "lab_owner", "analysis", "enterococci_cfu_per_100ml", "sample_number"),
           col_types = c("date", "text", "text", "text", "text", "numeric"),
           skip = 1)

# Have a look at the data
glimpse(lims)
summary(lims)
head(lims)
tail(lims)
slice_sample(lims, n = 50)
```

```{r clean-site-codes}
lims |>
  distinct(site_id) |>
  arrange(site_id)

# Clean the codes
lims <- lims |>
  mutate(site_id = case_when(
    site_id == "CN9_EAST" ~ "CS09",
    .default = str_to_upper(site_id)
  ))
```

```{r clean-data}
# Which sites in the LIMS data are not in the database or in the Excel data?
sites_not_in <- dplyr::setdiff(
  lims |> mutate(site_id = str_to_upper(site_id)) |> distinct(site_id),
  tbl(con, I("coastal.sites")) |> collect() |> distinct(site_id)
) |> pull()

# Current sites
tbl(con, I("coastal.sites")) |> collect()

# Filter LIMS data for only sites that are currently being monitored
lims |>
  filter((site_id %in% sites_not_in)) |>
  group_by(site_id) |>
  summarise(
    min_date = min(sample_date),
    max_date = max(sample_date),
    sample_size = n(),
    min_value = min(enterococci_cfu_per_100ml),
    max_value = max(enterococci_cfu_per_100ml)
            ) |>
  filter(sample_size > 30)

# Which of these have been running for a long time?
# If we confirm that we want to include these data, I will need a mapping
# For now, I exclude the sites that are not currently monitored
```

```{r clean-values}
lims |>
  distinct(enterococci_cfu_per_100ml)

lims <- lims |>
  mutate(enterococci_cfu_per_100ml = case_when(
    enterococci_cfu_per_100ml == 450 ~ ">150",
    enterococci_cfu_per_100ml == 2419 ~ ">2419",
    enterococci_cfu_per_100ml == 4500 ~ ">1500",
    enterococci_cfu_per_100ml == 24196 ~ ">24196",
    enterococci_cfu_per_100ml |> str_detect("nr|og") ~ NA,
    .default = enterococci_cfu_per_100ml
  )) |>
  filter(!is.na(enterococci_cfu_per_100ml))
  
# Create new numeric values from character data
lims <- lims |>
  filter(!is.na(enterococci_cfu_per_100ml)) |>
  mutate(enterococci_cfu_per_100ml_numeric = case_when(
    str_detect(enterococci_cfu_per_100ml, "<") ~ str_remove(enterococci_cfu_per_100ml, "<") |> as.numeric() / 3,
    str_detect(enterococci_cfu_per_100ml, ">") ~ str_remove(enterococci_cfu_per_100ml, ">") |> as.numeric() * 3,
    .default = as.numeric(enterococci_cfu_per_100ml)
  ))

# Write to csv
lims |>
  mutate(sample_date = date(sample_date)) |>
  rename(censored_value = enterococci_cfu_per_100ml, numeric_value = enterococci_cfu_per_100ml_numeric) |>
  select(sample_date, site_id, lab_owner, analysis, censored_value, numeric_value, sample_number) |>
  write_csv("02 Output Data/01 CSV/lims_data_clean.csv")
```

Can CN05A be connected to CN05 (Green Point Pump Station)?
Can CN06A be connected to CN06 Three Anchor Bay NW Rocks or CN06C Rocklands Beach?
CN12B can be connected to CN12A because they are combined in the spreadsheets
CN15?
CN20O Maiden's Cove Tidal Pool Outside?
CS01 connected to CS01A Kalk Bay Harbour Beach?
CS05 connected to what?
CS08 connected to what?
CS12 connected to what?
CS24E and CS24W connected to what?
XCS16
XCS33

# Blue Flag Data

```{r read-blue-flag-data}
path <- "../../Infinity/Infinity Projects - Files/190-56 Coastal WQ/05_Data/03 Water Quality/Coastal Water Quality/Blue Flag Data/"
filename <- str_c(path, list.files(path), sep = "")

blue_flag_files <- filename |>
  map_df(~tibble(
    filename = .x,
    sheet = excel_sheets(.x)
    ))

# Read in data
n = 40
(file <- blue_flag_files |> slice(n) |> pull(1))
(sheet <- blue_flag_files |> slice(n) |> pull(2))
(blue_flag_append <- read_excel(
  path = file, 
  sheet = sheet, 
  range = "B30:D130",
  col_names = c("sample_date", "ecoli_cfu_per_100ml", "faecal_streptococci_cfu_per_100ml"),
  col_types = c("date", "text", "text")
  ) |>
  filter(!if_all(1:3, is.na)) |>
  mutate(
    file = file |> str_remove(path),
    sheet = sheet))

(blue_flag_data <- bind_rows(blue_flag_data, blue_flag_append))

# Check Blue Flag Data
blue_flag_data

# Write Blue Flag data to file
write_csv(blue_flag_data, "02 Output Data/blue_flag_raw.csv")

# Read Blue Flag data
blue_flag_data <- read_csv("02 Output Data/blue_flag_raw.csv")
```

```{r clean-blue-flag}
# Check sites
blue_flag_data |>
  distinct(sheet) |>
  arrange(sheet)

blue_flag_data |>
  arrange(-as.integer(faecal_streptococci_cfu_per_100ml))

blue_flag_data |>
  filter(if_any(everything(), is.na))

blue_flag <- blue_flag_data |>
  mutate(site_name = case_when(
    sheet |> str_to_lower() |> str_detect("bikini") ~ "Bikini Beach",
    sheet |> str_to_lower() |> str_detect("clifton") ~ "Clifton Fourth",
    sheet |> str_to_lower() |> str_detect("fish") ~ "Fish Hoek",
    sheet |> str_to_lower() |> str_detect("llan") ~ "Llandudno",
    sheet |> str_to_lower() |> str_detect("melkbos") ~ "Melkbos",
    sheet |> str_to_lower() |> str_detect("muiz") ~ "Muizenberg",
    sheet |> str_to_lower() |> str_detect("seafo") ~ "Seaforth",
    sheet |> str_to_lower() |> str_detect("silw") ~ "Silwerstroom",
    sheet |> str_to_lower() |> str_detect("strandf") ~ "Strandfontein",
    .default = sheet
  )) |>
  select(site_name, sample_date, faecal_streptococci_cfu_per_100ml)

blue_flag <- blue_flag |>
  filter(!if_any(everything(), is.na))
```

```{r check-blue-flag}
# Check the highest values
blue_flag |>
  arrange(-as.integer(faecal_streptococci_cfu_per_100ml))

# Have a look at the data
blue_flag |>
  mutate(faecal_streptococci_cfu_per_100ml = faecal_streptococci_cfu_per_100ml |> as.numeric()) |>
  filter(faecal_streptococci_cfu_per_100ml < 400) |>
  ggplot(aes(x = faecal_streptococci_cfu_per_100ml |> as.numeric(), fill = site_name)) +
  geom_histogram(position = position_dodge())

# Check that everything can be converted to an integer
blue_flag |>
  mutate(faecal_streptococci_cfu_per_100ml = as.numeric(faecal_streptococci_cfu_per_100ml)) |>
  filter(is.na(faecal_streptococci_cfu_per_100ml))

# Add censored values
blue_flag <- blue_flag |>
  mutate(censored_value = case_when(
    faecal_streptococci_cfu_per_100ml %in% c(150, 151) ~ ">150",
    faecal_streptococci_cfu_per_100ml == 300 ~ ">300",
    faecal_streptococci_cfu_per_100ml == 3000 ~ ">3000",
    .default = as.character(faecal_streptococci_cfu_per_100ml)
    )) |>
  select(site_name, sample_date, censored_value)

# Export the clean data
write_csv(blue_flag, "02 Output Data/blue_flag_clean.csv")
```

# Daily Samples

## Extract data from pdf

```{r extract-pdf}
# Select the pdf that you want to import
path <- "../../Infinity/Coastal Water Quality - Documents/Shared Results/Strand Daily/"
files_list <- list.files("../../Infinity/Coastal Water Quality - Documents/Shared Results/Strand Daily/", pattern = ".*WALAB.*.pdf")
files <- str_c(path, files_list, sep = "")
files

read_walab_daily_data <- function(filename){
  # Extract the pdf data
  pdf_data <- pdf_text(filename) |>
    str_split("\n")
  
  # Extract the sample date
  sample_date <- pdf_data |>
    unlist() |> 
    keep(~str_detect(.x, "Sample taken")) |>
    str_extract("\\d{1,2}\\s[:alpha:]{3,12}\\s\\d{4}") |>
    pluck(1) |>
    dmy()
  
  # Extract the analysis date
  analysis_completed <- pdf_data |>
    unlist() |> 
    keep(~str_detect(.x, "Analysis completed")) |>
    str_extract("\\d{1,2}\\s[:alpha:]{3,12}\\s\\d{4}") |>
    pluck(1) |>
    dmy()
  
  # Extract the sample temperature
  sample_temp_c <- pdf_data |>
    unlist() |> 
    keep(~str_detect(.x, "Sample temperature:")) |>
    str_extract("(?<=(Sample temperature:\\s))\\d{1,2}\\.\\d") |>
    pluck(1) |>
    as.numeric()
  
  # Extract tabular data
  extract_table <- function(page){
    page |>
      unlist() |> 
      head_while(~!str_detect(.x, "colony forming units")) |>
      tail_while(~!str_detect(.x, "Sample number")) |>
      discard(~.x == "") |>
      unlist()
  }
  
  # Extract table from all pages of pdf
  table <- map(pdf_data, ~extract_table(.x)) |>
    unlist()
  
  # Extract sample number
  sample_number <- table |>
    str_sub(5,15) |>
    str_trim()
  
  # Extract sample text
  sample_text <- table |>
    str_sub(19,95) |>
    str_trim()
  
  # Extract result
  sample_result <- table |>
    str_sub(90) |>
    str_trim()
  
  # Make a table with the extracted data
  tbl <- tibble(number = sample_number,
         text = sample_text,
         result = sample_result) |> 
    mutate(number = if_else(number == "", NA, number)) |>
    fill(number, .direction = "down") |> 
    group_by(number) |>
    summarise(text = str_c(text, collapse = ""), result = first(result)) |>
    select(text, result) |>
    mutate(
      site_id = case_when(
        text |> str_to_lower() |> str_detect("cape|sands") ~ "CS30",
        text |> str_to_lower() |> str_detect("blue|building") ~ "CS31",
        .default = text
      ),
      sample_time = str_extract(text, "\\d{1,2}h\\d{2}") |> str_replace("h", ":"),
      sample_date = sample_date,
      analysis_completed = analysis_completed,
      sample_temp_c = sample_temp_c,
      filename = filename |> str_remove(path)
      ) |>
    select(-text) |>
    rename(enterococci_cfu_per_100ml = result) |>
    mutate(site_id = case_when(
      str_detect(site_id, "\\d{3}") & str_detect(site_id, "1$") ~ str_replace(site_id, "1$", "I"),
      str_detect(site_id, "\\d{3}") & str_detect(site_id, "0$") ~ str_replace(site_id, "0$", "O"),
      .default = site_id
    )) |>
    mutate(enterococci_cfu_per_100ml = str_replace(enterococci_cfu_per_100ml, "None found", "ND"))
}

files |>
  keep_at(10) |>
  map_df(~read_walab_daily_data(.x))

read_walab_daily_data("../../Infinity/Coastal Water Quality - Documents/Shared Results/Strand Daily/2024-09-23 Strand WALAB.pdf")

daily_data <- files |>
  map_df(~read_walab_daily_data(.x))

daily_data <- daily_data |>
  filter(if_any(everything(), is.na))

# Add data that could not be added automatically
daily_data <- daily_data |>
  add_case(enterococci_cfu_per_100ml = "39", site_id = "ICS12", sample_time = "10:13", sample_date = ymd("2025-02-20"), analysis_completed = ymd("2025-02-23"), sample_temp_c = 2.8, filename = "2025-02-20_WALAB_Atlantic.pdf") |>
  add_case(enterococci_cfu_per_100ml = "5", site_id = "ICS10", sample_time = "10:19", sample_date = ymd("2025-02-20"), analysis_completed = ymd("2025-02-23"), sample_temp_c = 2.8, filename = "2025-02-20_WALAB_Atlantic.pdf") |>
  add_case(enterococci_cfu_per_100ml = "3", site_id = "CN31", sample_time = "10:50", sample_date = ymd("2025-02-20"), analysis_completed = ymd("2025-02-23"), sample_temp_c = 2.8, filename = "2025-02-20_WALAB_Atlantic.pdf") |>
  add_case(enterococci_cfu_per_100ml = "1", site_id = "ICS12", sample_time = "11:15", sample_date = ymd("2025-01-15"), analysis_completed = ymd("2025-01-17"), sample_temp_c = 2.8, filename = "2025-01-15_WALAB_Atlantic.pdf") |>
  add_case(enterococci_cfu_per_100ml = "4", site_id = "ICS10", sample_time = "11:18", sample_date = ymd("2025-01-15"), analysis_completed = ymd("2025-01-17"), sample_temp_c = 2.8, filename = "2025-01-15_WALAB_Atlantic.pdf") |>
  add_case(enterococci_cfu_per_100ml = "42", site_id = "CN31", sample_time = "10:45", sample_date = ymd("2025-01-15"), analysis_completed = ymd("2025-01-17"), sample_temp_c = 2.8, filename = "2025-01-15_WALAB_Atlantic.pdf")

# Clean daily data
sites_capture <- daily_data |>
  distinct(filename) |>
  pull()

files |> str_remove(path) |>
setdiff(sites_capture)

# Check the values in the data
daily_data <- daily_data |>
  mutate(enterococci_cfu_per_100ml = str_remove(enterococci_cfu_per_100ml, "\\s")) |>
  mutate(enterococci_cfu_per_100ml = str_remove(enterococci_cfu_per_100ml, "`")) |>
  arrange(enterococci_cfu_per_100ml)

# Check for duplicates
daily_data |>
  group_by(site_id, sample_date, enterococci_cfu_per_100ml) |>
  filter(n() > 1)

# Write to file
daily_data |>
  write_csv("02 Output Data/daily_clean_strand.csv")

# Check summary of daily data
summary(daily_data)

# Copy to database
copy_to(con, tbl, name = "coastal", overwrite = TRUE)
```

```{r clean-atlantic-daily}
read_csv("../02 Output Data/daily_clean_atlantic.csv") |>
  mutate(site_id = if_else(site_id == "ISC13", "HB13", site_id)) |>
  write_csv("../02 Output Data/daily_clean_atlantic.csv")
```

# SABS Daily Data

```{r sabs-daily-data}
# Read in data
sabs_daily <- read_excel("../../Infinity/Coastal Water Quality - Documents/Shared Results/Atlantic Daily/Split Camps Bay daily.xlsx")

# Clean data
sabs_daily <- sabs_daily |>
  rename(
    sample_date = `Date Sampled`,
    date_received = `Date Received &Analysed`,
    site_description = `Site_Description`,
    site_id = `Site_Code`,
    sample_time = Time,
    enterococci_cfu_per_100ml = `Enterococci cfu/100ml`,
    ecoli_cfu_per_100ml = `E.coli cfu/100ml`,
    issues = Issues,
    comments = `Issues comments`
  ) |>
  mutate(
    sample_date = dmy(sample_date),
    date_received = dmy(date_received),
    sample_time = str_replace(sample_time, "h", ":") |> str_c(":00", sep = "") |> hms::as_hms(),
    enterococci_cfu_per_100ml = str_replace(enterococci_cfu_per_100ml, "±", "") |> str_replace("None found", "ND")
  ) |>
  filter(sample_date < "2024-10-05") |>
  select(1:6)

# Write SABS data to file
write_csv(sabs_daily, "02 Output Data/sabs_daily_clean.csv")
```

