---
title: "historical_coastal_water_quality"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r import-packages, message = FALSE}
library(tidyverse)
library(dbplyr)
library(DBI)
library(pdftools)
library(readxl)
```

```{r database-connection}
# Enter credentials when prompted
con <- dbConnect(RPostgres::Postgres(),
  host = "aws-0-eu-central-1.pooler.supabase.com",
  port = 5432,
  user = str_c(rstudioapi::askForPassword(prompt = "Username: "), "wnrvdvesovnkmqbhkhjz", sep = "."),
  password = rstudioapi::askForPassword(),
  dbname = "postgres"
)
```

```{r prepare-city-colours}
city_blue <- "#0098c5"
city_green <- "#bacf00"
city_pink <- "#c8006f"
city_teal <- "#005870"
city_dark_green <- "#446414"
city_red <- "#9d2235"
city_brown <- "#47292e"
city_tan <- "#98871f"
```

```{r data-import}
# Import the results and sites data from database
results <- tbl(con, I("coastal.results_view")) |> collect()

# Sort results
results_sorted <- results %>%
  arrange(sample_date)
view(results_sorted)

# Check frequency of sampling
results_frequency <- results %>%
  group_by(site_id, monitoring_group) %>%
  summarize(
    n_samples = n_distinct(sample_date),
    min_date = min(sample_date),
    max_date = max(sample_date)
  ) %>%
  mutate(
    date_range = as.numeric(max_date - min_date),
    frequency = ifelse(n_samples > 1, date_range / (n_samples - 1), 0) # Calculate average interval
  )
```

# Structure of our data

- How many records?
- At what frequency was data collected?
- How many records for each site?
- For what periods do we have data?
- What is the source of the data for the different periods?
- Is there any missing data?
- Is there any data that we still need to get hold of?

# Data distribution

## Overall distribution

## Overall distribution by coastline

## Overall distribution over time

# Hazen categories over time

- How have the categories for each site changed over time?
- Maps of water quality with sites coloured by category

# Percentage exceedances over time

- Maps of water quality with points sized by percentage exceedances

# Distribution of the data by year for each site

- Which sites are suitable for this analysis?

```{r Violin-plots-yearly}
library(lubridate)

results_hazen <- results_sorted %>%
  group_by(site_id, year = year(sample_date)) %>%
  summarise(
    min_date = min(sample_date, na.rm = TRUE),
    max_date = max(sample_date, na.rm = TRUE),
    n_enterococci = sum(!is.na(numeric_value)),
    hazen95 = quantile(numeric_value, 0.95, type = 5, na.rm = TRUE),
    hazen90 = quantile(numeric_value, 0.90, type = 5, na.rm = TRUE),
  hazen_category_enterococci = case_when(
      n_enterococci < 10 ~ "Not enough samples",
      hazen95 <= 100 ~ "Excellent",
      hazen95 <= 200 ~ "Good",
      hazen95 > 200 & hazen90 > 185 ~ "Poor",
      hazen95 > 200 & hazen90 < 185 ~ "Sufficient",
      TRUE ~ NA_character_
  ))

#Merge categories back into the original data frame
violin_data <- results_sorted %>%
  mutate(year = year(sample_date)) %>%
  left_join(results_hazen, by = c("site_id", "year"))

# Get unique site IDs (optional, for reference)
unique_sites <- unique(violin_data$site_id)
print(paste("Available Site IDs:", paste(unique_sites, collapse = ", ")))

# Prompt the user for a site ID
selected_site <- readline(prompt = "Enter the Site ID for the plot: ")

# Filter data for the selected site
site_violin_data <- violin_data %>%
  filter(site_id == selected_site)

# Check if the selected site exists in the data
if (nrow(site_violin_data) > 0) {
  # Create the violin plot for the selected site
  plot <- ggplot(site_violin_data, aes(x = factor(year), y = numeric_value, fill = hazen_category_enterococci)) +
    geom_violin(trim = TRUE) +
    geom_point(aes(group = factor(year)), # Group points within each violin
               position = position_jitter(width = 0.2),
               alpha = 0.3,
               size = 1) +
    scale_y_log10() +
    labs(
      title = paste("Enterococci Distribution by Year - Site:", selected_site),
      x = "Year",
      y = "Enterococci (log10 scale)",
      fill = "Hazen Category"
    ) +
    theme_minimal() +
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
          axis.ticks.length = unit(-0.15, "cm"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))

  # Print the plot for the selected site
  print(plot)
} else {
  cat(paste("Error: Site ID '", selected_site, "' not found in the data.\n"))
}

```

# Seasonal trends in water quality across years

- How do the data change over the year?
```{r Seasonal-Climatology-monthly}

#Calculate Hazen Categories based on monthly data
results_hazen_monthly <- results_sorted %>%
  group_by(site_id, month = month(sample_date, label = TRUE)) %>%
  summarise(
    min_date = min(sample_date, na.rm = TRUE),
    max_date = max(sample_date, na.rm = TRUE),
    n_samples = sum(!is.na(numeric_value)),
    hazen95 = quantile(numeric_value, 0.95, type = 5, na.rm = TRUE),
    hazen90 = quantile(numeric_value, 0.90, type = 5, na.rm = TRUE),
    hazen_category = case_when(
      n_samples < 10 ~ "Not enough samples",
      hazen95 <= 10 ~ "Excellent",
      hazen95 <= 20 ~ "Good",
      hazen95 > 20 & hazen90 > 18 ~ "Poor",
      hazen95 > 20 & hazen90 < 18 ~ "Sufficient",
      TRUE ~ NA_character_
    )
  )

#Merge with original data
violin_data_monthly <- results_sorted %>%
  mutate(month = month(sample_date, label = TRUE)) %>%
  left_join(results_hazen_monthly, by = c("site_id", "month"))

# Get unique site IDs (optional, for reference)
unique_sites <- unique(violin_data_monthly$site_id)
print(paste("Available Site IDs:", paste(unique_sites, collapse = ", ")))

# Prompt the user for a site ID
selected_site <- readline(prompt = "Enter the Site ID for the plot: ")

# Filter data for the selected site
site_violin_data <- violin_data_monthly %>%
  filter(site_id == selected_site)

# Check if the selected site exists in the data
if (nrow(site_violin_data) > 0) {
  # Create the violin plot for the selected site
 plot <- ggplot(site_violin_data, aes(x = month, y = numeric_value, fill = hazen_category)) +
    geom_violin(trim = TRUE) +
    geom_point(aes(group = month), # Group points within each violin
               position = position_jitter(width = 0.2),
               alpha = 0.3,
               size = 1) +
    scale_y_log10() + 
    labs(
      title = paste("Monthly Climatology of Enterococci distribution - Site:", selected_site),
      x = "Month",
      y = "Enterococci (log10 scale)",
      fill = "Hazen Category"
    ) +
    theme_minimal() +
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
          axis.ticks.length = unit(-0.15, "cm"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))

 # Print the plot for the selected site
  print(plot)
} else {
  cat(paste("Error: Site ID '", selected_site, "' not found in the data.\n"))
}
```

# Seasonal trends in water quality across every year
```{r Seasonality-across-every-yr, fig.height = 10}

#Filter site CN11
selected_site <- "CN11"
results_hazen_monthly2 <- results_sorted %>%
  filter(site_id == selected_site)

#Calculate Hazen Categories based on monthly data
results_hazen_monthly2_1 <- results_hazen_monthly2 %>%
  mutate(year = year(sample_date),
         month = month(sample_date, label = TRUE)) %>%
  group_by(site_id, year, month) %>%
  summarise(
    min_date = min(sample_date, na.rm = TRUE),
    max_date = max(sample_date, na.rm = TRUE),
    n_samples = sum(!is.na(numeric_value)),
    hazen95 = quantile(numeric_value, 0.95, type = 5, na.rm = TRUE),
    hazen90 = quantile(numeric_value, 0.90, type = 5, na.rm = TRUE),
    hazen_category = case_when(
      n_samples < 10 ~ "Not enough samples",
      hazen95 <= 10 ~ "Excellent",
      hazen95 <= 20 ~ "Good",
      hazen95 > 20 & hazen90 > 18 ~ "Poor",
      hazen95 > 20 & hazen90 < 18 ~ "Sufficient",
      TRUE ~ NA_character_
    ),
    .groups = 'drop'
  )

#Merge with original data
violin_data_monthly2 <- results_hazen_monthly2 %>%
  mutate(year = year(sample_date),
    month = month(sample_date, label = TRUE)) %>%
  left_join(results_hazen_monthly2_1, by = c("site_id", "year", "month"))

# Get the unique years in your data for the selected site
unique_years <- unique(violin_data_monthly2$year)
num_years <- length(unique_years)

if (num_years > 0) {
  # Split the years into two roughly equal groups
  mid_point <- ceiling(num_years / 2)
  years_group1 <- sort(unique_years)[1:mid_point]
  years_group2 <- sort(unique_years)[(mid_point + 1):num_years]

  # --- Plot 1: First group of years ---
  plot_group1 <- violin_data_monthly2 %>%
    filter(year %in% years_group1) %>%
    ggplot(aes(x = month, y = numeric_value, fill = hazen_category)) +
    geom_violin(trim = TRUE, width = 0.8) +
    geom_point(aes(group = month),
               position = position_jitter(width = 0.2),
               alpha = 0.6,
               size = 0.8) +
    scale_y_log10() +
    facet_wrap(~ year, ncol = 2, strip.position = "right") + 
    labs(
      title = paste("Monthly Seasonality of Enterococci - Site:", selected_site, "(Years", min(years_group1), "-", max(years_group1), ")"),
      x = "Month",
      y = "Enterococci (log10 scale)",
      fill = "Hazen Category") +
    theme_minimal() +
    theme(
    axis.line = element_line(colour = "black"),
    axis.ticks.y = element_line(colour = "black"),
    axis.ticks.length.y = unit(-0.15, "cm"),
    axis.text.y = element_text(colour = "black", size = 10, margin = margin(r = 3)),
    axis.title.y = element_text(colour = "black", size = 12, margin = margin(r = 5)),
    axis.line.x = element_line(colour = "black"),
    axis.ticks.x = element_line(colour = "black"),
    axis.ticks.length.x = unit(-0.15, "cm"),
    axis.text.x = element_text(colour = "black", size = 10, angle = 45, hjust = 1),
    axis.title.x = element_text(colour = "black", size = 12, margin = margin(t = 5)),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),
    strip.text.y.right = element_text(colour = "black", size = 10, angle = 0, vjust = 0.8, hjust = -0.2, margin = margin(l = 0)),
    strip.background = element_rect(colour = NA, fill = NA),
    strip.placement = "inside",
    plot.title = element_text(colour = "black", size = 14),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")
    )
  print(plot_group1)

  # --- Plot 2: Second group of years (if there are more than one year) ---
  if (length(years_group2) > 0) {
    plot_group2 <- violin_data_monthly2 %>%
      filter(year %in% years_group2) %>%
      ggplot(aes(x = month, y = numeric_value, fill = hazen_category)) +
      geom_violin(trim = TRUE, width = 0.8) +
      geom_point(aes(group = month),
                 position = position_jitter(width = 0.2),
                 alpha = 0.6,
                 size = 0.8) +
      scale_y_log10() +
      facet_wrap(~ year, ncol = 2, strip.position = "right") +
      labs(
        title = paste("Monthly Seasonality of Enterococci - Site:", selected_site, "(Years", min(years_group2), "-", max(years_group2), ")"),
        x = "Month",
        y = "Enterococci (log10 scale)",
        fill = "Hazen Category"
      ) +
      theme_minimal() +
      theme(
    axis.line = element_line(colour = "black"),
    axis.ticks.y = element_line(colour = "black"),
    axis.ticks.length.y = unit(-0.15, "cm"),
    axis.text.y = element_text(colour = "black", size = 10, margin = margin(r = 5)),
    axis.title.y = element_text(colour = "black", size = 12, margin = margin(r = 10)),
    axis.line.x = element_line(colour = "black"),
    axis.ticks.x = element_line(colour = "black"),
    axis.ticks.length.x = unit(-0.15, "cm"),
    axis.text.x = element_text(colour = "black", size = 10, angle = 45, hjust = 1),
    axis.title.x = element_text(colour = "black", size = 12, margin = margin(t = 10)),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),
    strip.text.y.right = element_text(colour = "black", size = 10, angle = 0, vjust = 0.8, hjust = -0.2, margin = margin(l = 0)),
    strip.background = element_rect(colour = NA, fill = NA),
    strip.placement = "inside",
    plot.title = element_text(colour = "black", size = 14)
      )
    print(plot_group2)
  }
} else {
  print("No data available for the selected site.")
}

```


# Time-series trends in water quality data

- Changes in Hazen 95th and Hazen 90th over time for each site

```{R Monthly routine monitoring group}
# Filter routine monitoring group
routine_data <- results_sorted %>%
  filter(monitoring_group == "routine")
unique(routine_data$numeric_value)

monthly_routine <- routine_data %>%
  mutate(
    year_month = format(sample_date, "%Y-%m") 
  ) %>%
  group_by(site_id, year_month, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(monthly_routine)

#Plotting
#Monthly plots for each site are too cluttered, would prefer visualization on a yearly scale
library(ggplot2)

# Define colour palette for the labs
lab_colours <- c("SABS" = "red", "SSVC" = "blue", "WLAB" = "green")

unique_sites1 <- unique(monthly_routine$site_id)

for (site in unique_sites1) {
  site_data <- monthly_routine %>%
    filter(site_id == site) %>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year_month, y = enterococci_value, group = interaction(lab_code, percentile), colour = 
    lab_code, 
    linetype =percentile)) +
    geom_line(linewidth = 0.7)+
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    scale_linetype_manual(values = c("p90_enterococci" = "solid", "p95_enterococci" = "dashed"),
                          labels = c("90th Percentile", "95th Percentile")) +
    labs(
      title = paste("Routine Monitoring Monthly Enterococci Percentiles - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

```{R Yearly routine monitoring group}

yearly_routine <- routine_data %>%
  mutate(
    year = format(sample_date, "%Y")
  ) %>%
  group_by(site_id, year, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(yearly_routine)

#Plotting
unique_sites2 <- unique(yearly_routine$site_id)

for (site in unique_sites2) {
  site_data <- yearly_routine %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year, y = enterococci_value, group = interaction(lab_code, percentile), colour = lab_code, 
    linetype =percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    scale_linetype_manual(values = c("p90_enterococci" = "solid", "p95_enterococci" = "dashed"),
                          labels = c("90th Percentile", "95th Percentile")) +
    labs(
      title = paste("Routine Monitoring Yearly Enterococci Percentiles - Site:", site),
      x = "Year",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot)
}
```

```{R Monthly scale daily monitoring group}
# Filter daily monitoring group
daily_monitoring <- results_sorted %>%
  filter(monitoring_group == "daily")
unique(daily_monitoring$numeric_value)

monthly_daily_monitoring <- daily_monitoring %>%
  mutate(
    year_month = format(sample_date, "%Y-%m")
  ) %>%
  group_by(site_id, year_month, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(monthly_daily_monitoring)

#Plotting
unique_sites3 <- unique(monthly_daily_monitoring$site_id)

for (site in unique_sites3) {
  site_data <- monthly_daily_monitoring %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year_month, y = enterococci_value, group = interaction(lab_code, percentile), colour = 
    lab_code, 
    linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    labs(
      title = paste("Monthly Enterococci Percentiles for daily monitoring - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

```{R Daily scale daily monitoring group}

daily_percentiles_daily_mon <- daily_monitoring %>%
  group_by(site_id, sample_date, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(daily_percentiles_daily_mon)
#Don't think its necessary to plot the daily monitoring data as yearly averages because we only have 2 years worth of data

#Plotting
#90th & 95th Percentiles are equal; only 1 unique data point for each site, for each lab, for each day
unique_sites4 <- unique(daily_percentiles_daily_mon$site_id)
library(scales) # For date formatting

for (site in unique_sites4) {
  site_data <- daily_percentiles_daily_mon %>%
    filter(site_id == site) %>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = sample_date, y = enterococci_value, group = interaction(lab_code, percentile), colour = 
    lab_code, 
    linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    scale_linetype_manual(values = c("p90_enterococci" = "solid", "p95_enterococci" = "dashed"),
                          labels = c("90th Percentile", "95th Percentile")) +
    scale_x_date(
      labels = date_format("%Y-%m-%d"),
      breaks = "2 weeks"             
    ) +
    labs(
      title = paste("Daily Enterococci Percentiles for daily monitoring - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

```{R Monthly Blue flag}
# Filter blue flag group
blue_flag <- results_sorted %>%
  filter(monitoring_group == "blue_flag")
unique(blue_flag$numeric_value)

monthly_blue_flag <- blue_flag %>%
  group_by(site_id, sample_date, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(monthly_blue_flag)
#90th & 95th Hazen percentiles  are equal; every month of blue flag data has only one data point for each month, for each site and each lab

#Plotting
unique_sites5 <- unique(monthly_blue_flag$site_id)

for (site in unique_sites5) {
  site_data <- monthly_blue_flag %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = sample_date, y = enterococci_value, group = interaction(lab_code, percentile), colour =
    lab_code,linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values =lab_colours)+
    scale_x_date(
      labels = date_format("%Y-%m"),
      breaks = "4 months"             
    ) +
    labs(
      title = paste("Blue Flag Monthly Enterococci Percentiles - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

```{R Yearly Blue flag}

yearly_blue_flag <- blue_flag %>%
  mutate(
    year = format(sample_date, "%Y")
  ) %>%
  group_by(site_id, year, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(yearly_blue_flag)

#Plotting
unique_sites6 <- unique(yearly_blue_flag$site_id)

for (site in unique_sites6) {
  site_data <- yearly_blue_flag %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year, y = enterococci_value, group = interaction(lab_code, percentile), colour =
    lab_code,linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values =lab_colours)+
    labs(
      title = paste("Blue Flag Yearly Enterococci Percentiles - Site:", site),
      x = "Year",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
          axis.ticks.length = unit(-0.15, "cm"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

# Cumulative change in Hazen percentiles over time

- On net, has the site improved or deteriorated?

# Grouped analyses

## Group sites by region

## Group sites by features

- Stormwater drains
- River mouths
- Natural features (algal blooms)
- Marine outfalls