---
title: "historical_coastal_water_quality"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r import-packages, message = FALSE}
library(tidyverse)
library(dbplyr)
library(DBI)
library(pdftools)
library(readxl)
```

```{r database-connection}
# Enter credentials when prompted
con <- dbConnect(RPostgres::Postgres(),
  host = "aws-0-eu-central-1.pooler.supabase.com",
  port = 5432,
  user = str_c(rstudioapi::askForPassword(prompt = "Username: "), "wnrvdvesovnkmqbhkhjz", sep = "."),
  password = rstudioapi::askForPassword(),
  dbname = "postgres"
)
```

```{r prepare-city-colours}
city_blue <- "#0098c5"
city_green <- "#bacf00"
city_pink <- "#c8006f"
city_teal <- "#005870"
city_dark_green <- "#446414"
city_red <- "#9d2235"
city_brown <- "#47292e"
city_tan <- "#98871f"
```

```{r data-import}
# Import the results and sites data from database
results <- tbl(con, I("coastal.results_view")) |> collect()

# Sort results
results_sorted <- results %>%
  arrange(sample_date)
view(results_sorted)

# Check frequency of sampling
results_frequency <- results %>%
  group_by(site_id, monitoring_group) %>%
  summarize(
    n_samples = n_distinct(sample_date),
    min_date = min(sample_date),
    max_date = max(sample_date)
  ) %>%
  mutate(
    date_range = as.numeric(max_date - min_date),
    frequency = ifelse(n_samples > 1, date_range / (n_samples - 1), 0) # Calculate average interval
  )
```

# Structure of our data

- How many records?
- At what frequency was data collected?
- How many records for each site?
- For what periods do we have data?
- What is the source of the data for the different periods?
- Is there any missing data?
- Is there any data that we still need to get hold of?

# Data distribution

## Overall distribution

## Overall distribution by coastline

## Overall distribution over time

# Hazen categories over time

- How have the categories for each site changed over time?
- Maps of water quality with sites coloured by category

# Percentage exceedances over time

- Maps of water quality with points sized by percentage exceedances

# Distribution of the data by year for each site

- Which sites are suitable for this analysis?

# Seasonal trends in water quality across years

- How do the data change over the year?

# Time-series trends in water quality data

- Changes in Hazen 95th and Hazen 90th over time for each site

# Cumulative change in Hazen percentiles over time

- On net, has the site improved or deteriorated?

# Grouped analyses

## Group sites by region

## Group sites by features

- Stormwater drains
- River mouths
- Natural features (algal blooms)
- Marine outfalls

# Calculating routine monitoring group percentiles

## Monthly Scale

```{R Calculate monthly/yearly 90th & 95th Hazen percentile values of routine monitoring group}
# Filter routine monitoring group
routine_data <- results_sorted %>%
  filter(monitoring_group == "routine")
unique(routine_data$numeric_value)

monthly_routine <- routine_data %>%
  mutate(
    year_month = format(sample_date, "%Y-%m") 
  ) %>%
  group_by(site_id, year_month, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(monthly_routine)
```

## Yearly Scale

```{R Calculate monthly/yearly 90th & 95th Hazen percentile values of routine monitoring group}

yearly_routine <- routine_data %>%
  mutate(
    year = format(sample_date, "%Y")
  ) %>%
  group_by(site_id, year, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(yearly_routine)
```

#Plotting routine monitoring group percentiles

## Monthly Scale

```{R Plotting 90th & 95th Hazen percentiles of routine monitoring group}
#Monthly plots for each site are too cluttered, would prefer visualization on a yearly scale

library(ggplot2)

# Define colour palette for the labs
lab_colours <- c("SABS" = "red", "SSVC" = "blue", "WLAB" = "green")

unique_sites1 <- unique(monthly_routine$site_id)

for (site in unique_sites1) {
  site_data <- monthly_routine %>%
    filter(site_id == site) %>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year_month, y = enterococci_value, group = interaction(lab_code, percentile), colour = lab_code, 
    linetype =percentile)) +
    geom_line(linewidth = 0.7)+
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    scale_linetype_manual(values = c("p90_enterococci" = "solid", "p95_enterococci" = "dashed"),
                          labels = c("90th Percentile", "95th Percentile")) +
    labs(
      title = paste("Routine Monitoring Monthly Enterococci Percentiles - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

## Yearly Scale

```{R Plotting 90th & 95th Hazen percentiles of routine monitoring group}

unique_sites2 <- unique(yearly_routine$site_id)

for (site in unique_sites2) {
  site_data <- yearly_routine %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year, y = enterococci_value, group = interaction(lab_code, percentile), colour = lab_code, 
    linetype =percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    scale_linetype_manual(values = c("p90_enterococci" = "solid", "p95_enterococci" = "dashed"),
                          labels = c("90th Percentile", "95th Percentile")) +
    labs(
      title = paste("Routine Monitoring Yearly Enterococci Percentiles - Site:", site),
      x = "Year",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot)
}
```

# Calculating daily monitoring group percentiles

## Monthly Scale

```{R Calculate 90th & 95th Hazen percentiles for daily monitoring group}
# Filter daily monitoring group
daily_monitoring <- results_sorted %>%
  filter(monitoring_group == "daily")
unique(daily_monitoring$numeric_value)

monthly_daily_monitoring <- daily_monitoring %>%
  mutate(
    year_month = format(sample_date, "%Y-%m")
  ) %>%
  group_by(site_id, year_month, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(monthly_daily_monitoring)
```

## Daily Scale

```{R Calculate 90th & 95th Hazen percentiles for daily monitoring group}

daily_percentiles_daily_mon <- daily_monitoring %>%
  group_by(site_id, sample_date, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(daily_percentiles_daily_mon)
#Don't think its necessary to plot the daily monitoring data as yearly averages because we only have 2 years worth of data

```

#Plotting daily monitoring group percentiles

## Monthly Scale

```{r Plotting 90th & 95th Hazen percentiles for daily monitoring group}
unique_sites3 <- unique(monthly_daily_monitoring$site_id)

for (site in unique_sites3) {
  site_data <- monthly_daily_monitoring %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year_month, y = enterococci_value, group = interaction(lab_code, percentile), colour = lab_code, 
    linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    labs(
      title = paste("Monthly Enterococci Percentiles for daily monitoring - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

## Daily Scale

```{r Plotting 90th & 95th Hazen percentiles for daily monitoring group}
#90th & 95th Percentiles are equal; only 1 unique data point for each site, for each lab, for each day

unique_sites4 <- unique(daily_percentiles_daily_mon$site_id)
library(scales) # For date formatting

for (site in unique_sites4) {
  site_data <- daily_percentiles_daily_mon %>%
    filter(site_id == site) %>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = sample_date, y = enterococci_value, group = interaction(lab_code, percentile), colour = lab_code, 
    linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values = lab_colours) +
    scale_linetype_manual(values = c("p90_enterococci" = "solid", "p95_enterococci" = "dashed"),
                          labels = c("90th Percentile", "95th Percentile")) +
    scale_x_date(
      labels = date_format("%Y-%m-%d"),
      breaks = "2 weeks"             
    ) +
    labs(
      title = paste("Daily Enterococci Percentiles for daily monitoring - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(
      axis.line = element_line(colour = "black"),
      axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

# Calculating blue flag percentiles

## Monthly Scale

```{R Calculating Blue flag 90th & 95th Hazen percentiles}
# Filter blue flag group
blue_flag <- results_sorted %>%
  filter(monitoring_group == "blue_flag")
unique(blue_flag$numeric_value)

monthly_blue_flag <- blue_flag %>%
  group_by(site_id, sample_date, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(monthly_blue_flag)
#90th & 95th Hazen percentiles  are equal; every month of blue flag data has only one data point for each month, for each site and each lab
```

## Yearly Scale

```{R Calculating Blue flag 90th & 95th Hazen percentiles}

yearly_blue_flag <- blue_flag %>%
  mutate(
    year = format(sample_date, "%Y")
  ) %>%
  group_by(site_id, year, lab_code) %>%
  summarize(
    p90_enterococci = round(quantile(numeric_value, probs = 0.90, na.rm = TRUE, type = 5), 0),
    p95_enterococci = round(quantile(numeric_value, probs = 0.95, na.rm = TRUE, type = 5), 0),
    .groups = "drop"
  )
view(yearly_blue_flag)
```

# Plotting blue flag percentiles

## Monthly Scale

```{R Plotting Blue flag 90th & 95th Hazen percentiles}
unique_sites5 <- unique(monthly_blue_flag$site_id)

for (site in unique_sites5) {
  site_data <- monthly_blue_flag %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = sample_date, y = enterococci_value, group = interaction(lab_code, percentile), colour =
    lab_code,linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values =lab_colours)+
    scale_x_date(
      labels = date_format("%Y-%m"),
      breaks = "4 months"             
    ) +
    labs(
      title = paste("Blue Flag Monthly Enterococci Percentiles - Site:", site),
      x = "Sampling date",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
      axis.ticks.length = unit(-0.15, "cm"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

## Yearly Scale

```{R Plotting Blue flag 90th & 95th Hazen percentiles}

unique_sites6 <- unique(yearly_blue_flag$site_id)

for (site in unique_sites6) {
  site_data <- yearly_blue_flag %>%
    filter(site_id == site)%>%
    pivot_longer(cols = c(p90_enterococci, p95_enterococci),
                 names_to = "percentile",
                 values_to = "enterococci_value")

  plot <- ggplot(site_data, aes(x = year, y = enterococci_value, group = interaction(lab_code, percentile), colour =
    lab_code,linetype = percentile)) +
    geom_line(linewidth = 0.7) +
    geom_point() +
    scale_colour_manual(values =lab_colours)+
    labs(
      title = paste("Blue Flag Yearly Enterococci Percentiles - Site:", site),
      x = "Year",
      y = "Enterococci (cfu/100ml)",
      colour = "Lab",
      linetype = "Percentile"
    ) +
    theme_minimal() +
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
          axis.ticks.length = unit(-0.15, "cm"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot) 
}
```

# Calculate the water quality category for each site for each year

1. Calculate the Hazen category for each site for each year (check with Francois, the code for the Hazen category is in Notebooks/hazen_calculation)
2. Plot a violin plot for each year (geom_violin) and colour the violins based on the water colour category (check with Francois)
3. Add the original points on top of the violin plots with alpha quite low (0.3) and the points scattered slightly (position_jitter())

```{r hazen-method}
library(lubridate)

results_hazen <- results_sorted %>%
  group_by(site_id, year = year(sample_date)) %>%
  summarise(
    min_date = min(sample_date, na.rm = TRUE),
    max_date = max(sample_date, na.rm = TRUE),
    n_enterococci = sum(!is.na(numeric_value)),
    hazen95 = quantile(numeric_value, 0.95, type = 5, na.rm = TRUE),
    hazen90 = quantile(numeric_value, 0.90, type = 5, na.rm = TRUE),
  hazen_category_enterococci = case_when(
      n_enterococci < 10 ~ "Not enough samples",
      hazen95 <= 100 ~ "Excellent",
      hazen95 <= 200 ~ "Good",
      hazen95 > 200 & hazen90 > 185 ~ "Poor",
      hazen95 > 200 & hazen90 < 185 ~ "Sufficient",
      TRUE ~ NA_character_
  ))
```
#  Violin plot, coloured by Hazen category, points jittered

```{r Violin plots}

#Merge categories back into the original data frame
violin_data <- results_sorted %>%
  mutate(year = year(sample_date)) %>%
  left_join(results_hazen, by = c("site_id", "year"))

# Get unique site IDs
unique_sites <- unique(violin_data$site_id)

for (site in unique_sites) {
  site_violin_data <- violin_data %>%
    filter(site_id == site)

plot <- ggplot(site_violin_data, aes(x = factor(year), y = numeric_value, fill = hazen_category_enterococci))+
    geom_violin(trim = TRUE) +
    geom_point(aes(group = factor(year)), # Group points within each violin
               position = position_jitter(width = 0.2), # Add slight horizontal jitter
               alpha = 0.3,
               size = 1) +
    scale_y_log10() +
    labs(
      title = paste("Enterococci Distribution by Year - Site:", site),
      x = "Year",
      y = "Enterococci (log10 scale)",
      fill = "Hazen Category"
    ) +
    theme_minimal() + 
    theme(axis.line = element_line(colour = "black"),
          axis.ticks = element_line(colour = "black"),
          axis.ticks.length = unit(-0.15, "cm"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))

  print(plot)
}
```
