---
title: "Recreational Node Summary Page"
author: "Francois van Schalkwyk"
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    reference_docx: recreational_node_overview_style_template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r load-packages}
library(tidyverse)
library(dbplyr)
library(DBI)
library(pdftools)
library(readxl)
library(extrafont)
library(sf)
library(basemaps)
library(raster)
library(patchwork)
library(units)
library(osmdata)
```

```{r db-connection}
# Connect to the Postgres database
# Enter credentials when prompted
con <- dbConnect(RPostgres::Postgres(),
  host = "aws-0-eu-central-1.pooler.supabase.com",
  port = 5432,
  user = str_c(rstudioapi::askForPassword(prompt = "Username: "), "wnrvdvesovnkmqbhkhjz", sep = "."),
  password = rstudioapi::askForPassword(),
  dbname = "postgres"
)
```

```{r prepare-city-colours}
city_blue <- "#0098c5"
city_green <- "#bacf00"
city_pink <- "#c8006f"
city_teal <- "#005870"
city_dark_green <- "#446414"
city_red <- "#9d2235"
city_brown <- "#47292e"
city_tan <- "#98871f"
```

# Create a map of all sites

```{r read-in-site-data}
sites_sf <- st_read(con, query = "SELECT * FROM coastal.sites", as_tibble = TRUE) |>
  st_cast("POINT") |>
  filter(active) |>
  arrange(site_description)
```

```{r}
# Read in coastline of Cape Town
cct <- st_read("../../Infinity/Infinity GIS - Server/base_data/sa/WC/cape_town/CPT Metropolitan/metropolitan municipality cpt.shp", as_tibble = TRUE) |>
  st_set_crs("EPSG: 4326")

-- temporary

cct <- st_read("../../Downloads/OneDrive_1_2025-04-22/metropolitan municipality cpt.shp", as_tibble = TRUE) |>
  st_set_crs("EPSG: 4326")

--
```

```{r city-overview, fig.cap = "Map of coastal water quality monitoring sites in Cape Town", fig.width = 6.4, fig.height = 9.8, dpi = 600}
# Plot city of Cape Town with sites
cct_sites <- cct |>
  ggplot() +
  geom_sf(colour = NA) +
  geom_sf(data = sites_sf, aes(colour = category, size = category)) +
  theme(
    legend.position = "bottom",
    legend.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.background = element_blank(),
    panel.background = element_blank(),
    text = element_text(family = "Century Gothic"),
    legend.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid = element_blank()
  ) +
  scale_colour_manual(values = c("Recreational Node" = city_green, "Coastal Monitoring Point" = "darkgrey")) +
  scale_size_manual(values = c("Recreational Node" = 1, "Coastal Monitoring Point" = 0.5))
```

# Create map of specific site

```{r plot-zoomed-data, fig.width = 6.4, fig.height = 9.8, dpi = 600}
# Select site_id
site <- "CN09"

# Choose a specific site to plot
sites_overview <- cct_sites +
  geom_sf(data = sites_sf |> filter(site_id == site), colour = city_pink, size = 5, shape = 16, alpha = 0.5) +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.background = element_rect(fill = city_teal)
  )
```

```{r build-basemap, fig.height = 6, fig.width=6, dpi = 600}
# Have a look at the types of basemaps
basemaps::get_maptypes()
set_defaults(map_service = "esri", map_type = "world_imagery")

# Create bounding box of rectangle
<<<<<<< HEAD
bbox <- sites |>
=======
bbox <- sites_sf |>
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
  filter(site_id == "CN09") |>
  st_transform("EPSG:3857") |>
  st_bbox()

# Set limits for bounding box
bbox <- bbox + c(-1000, -500, 1000, 500)

# Add sites to basemap
<<<<<<< HEAD
site <- sites |>
=======
site <- sites_sf |>
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
  filter(site_id == "CN09") |>
  st_transform("EPSG:3857")

# Build map
<<<<<<< HEAD
site_map <- basemap_ggplot(bbox, alpha = 0.7) +
=======
site_map <- basemap_ggplot(bbox) +
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
  geom_sf(data = site, alpha = 0.5, colour = city_pink, size = 5, shape = 16) +
  theme(
    panel.background = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_blank(),
    panel.grid = element_blank()
  )
```

# Inset the site map onto the overview map

```{r inset-map, fig.width = 6, fig.height = 6, dpi = 600}
<<<<<<< HEAD
sites_overview + inset_element(site_map, left = 0.4, right = 1, bottom = 0.4, top = 1)
=======
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
sites_overview / site_map
```

# Create a figure percentage of samples that have exceeded the threshold

```{r perc-exceedance-plot}
sites <- tbl(con, I("coastal.sites_view"))
results <- tbl(con, I("coastal.results_view"))

# Choose the site
<<<<<<< HEAD
site <- "CN22"
=======
site <- "CN09"
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872

# Join the tables
exceedances <- sites |>
  inner_join(results, by = "site_id") |>
<<<<<<< HEAD
  filter(site_id == "CN22", between(sample_date, "2024-01-01", "2025-01-01"), monitoring_group %in% c("routine", "daily")) |>
=======
  filter(site_id == "CN09", between(sample_date, "2024-01-01", "2025-01-01"), monitoring_group %in% c("routine", "daily")) |>
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
  collect() |>
  summarise(
    all_samples = n(),
    samples_exceed = sum(numeric_value > 240),
    samples_exceed_pct = round(sum(numeric_value > 240) / n() * 100, 0)
  ) |>
  mutate(
    label = str_c("of ", all_samples, " samples taken\nexceeded\nthe threshold\nvalue for\nEnterococci\nof 240 cfu per 100 mL", sep = "")
    )
```

```{r exceedances-figure}
# Create a figure for the percentage exceedances
<<<<<<< HEAD
ggplot(data = exceedances) +
=======
exceedance_plot <- ggplot(data = exceedances) +
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
  geom_text(aes(x = 2, y = 1, label = label), family = "Century Gothic", size = 10, hjust = 1, colour = "white") +
  geom_text(aes(x = 1, y = 1, label = str_c(samples_exceed_pct, "%", sep = ""), colour = samples_exceed_pct), family = "Century Gothic", size = 50, hjust = 0) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.background = element_rect(fill = city_teal),
    panel.grid = element_blank(),
    legend.position = "none"
  ) +
  scale_colour_distiller(type = "div", palette = "RdYlGn", limits = c(0, 50), direction = -1)

RColorBrewer::display.brewer.all()
```

# Hazen Categories

```{r}
hazen <- sites |>
  inner_join(results, by = "site_id") |>
  filter(site_id == "CN11", between(sample_date, "2024-01-01", "2025-01-01"), monitoring_group %in% c("routine", "daily")) |>
  collect() |>
  summarise(
    min_date = min(sample_date),
    max_date = max(sample_date),
    n = sum(!is.na(numeric_value)),
    hazen95 = quantile(numeric_value, 0.95, type = 5, na.rm = TRUE),
    hazen90 = quantile(numeric_value, 0.9, type = 5, na.rm = TRUE),
  hazen_category = case_when(
      n < 10 ~ "Not enough samples",
      hazen95 <= 100 ~ "Excellent",
      hazen95 <= 200 ~ "Good",
      hazen95 > 200 & hazen90 > 185 ~ "Poor",
      hazen95 > 200 & hazen90 < 185 ~ "Sufficient"
  ))
```

```{r hazen-figure}
  # Create a simple plot showing the category
<<<<<<< HEAD
  ggplot(data = hazen) +
=======
hazen_plot <- ggplot(data = hazen) +
>>>>>>> 71c6381fba5b4520ea94773c70471909f1f22872
  geom_rect(aes(xmin = 0.995, xmax = 1.005, ymin = 0.995, ymax = 1.005, fill = hazen_category)) +
  geom_text(aes(x = 1, y = 1, label = hazen_category), family = "Century Gothic", size = 35, hjust = 0.5, colour = "white") +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_fill_manual(name = "Hazen Water Quality Category", values = c("Poor" = "#A50026", "Sufficient" = "#A6D96A", "Good" = "#006837", "Excellent" = "#4575B4"))

RColorBrewer::display.brewer.all(type = "div")
RColorBrewer::brewer.pal(10, "RdYlGn")
RColorBrewer::brewer.pal(10, "RdYlBu")
RColorBrewer::brewer.pal(10, "RdBu")
```

# Sample frequency plots

```{r}
months <- seq(ymd("2024-01-01"), ymd("2025-01-01"), '1 month')

# Plot sample frequency in the reporting period
samples_plot <- sites |>
  inner_join(results, by = "site_id") |>
  filter(between(sample_date, "2024-01-01", "2024-12-31"), monitoring_group %in% c("routine", "daily", "blue_flag")) |>
  collect() |>
  filter(site_id == "CN09") |>
  group_by(month = floor_date(sample_date, "month"), site_id, monitoring_group) |>
  count() |>
  
  # Plot the data
  ggplot() +
  labs(title = "Number of samples taken per month") +
  geom_point(aes(x = month, y = n, colour = monitoring_group), position = position_dodge(preserve = "single"), size = 5) +
  geom_segment(aes(x = month, xend = month, y = 0, yend = n, colour = monitoring_group), size = 2) +
  
  # Scales
  scale_colour_manual(values = c("routine" = city_green, "daily" = city_pink, "blue_flag" = city_blue)) +
  scale_y_continuous(breaks = scales::breaks_pretty()) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  
  coord_cartesian(xlim = c(ymd("2024-01-01"), ymd("2024-12-31"))) +
  
# Plot styling
  theme(
    legend.position = "bottom",
    panel.background = element_blank(),
    axis.title = element_blank(),
    legend.title = element_blank(),
    text = element_text(family = "Century Gothic"),
    plot.background = element_rect(fill = city_teal),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.grid = element_blank(),
    plot.title = element_text(colour = "white", hjust = 0.5, face = "bold"),
    legend.background = element_blank(),
    legend.text = element_text(colour = "white")
  )
```

```{r results-plot}
results_plot <- sites |>
  inner_join(results, by = "site_id") |>
  filter(between(sample_date, "2024-01-01", "2024-12-31"), monitoring_group %in% c("routine", "daily", "blue_flag")) |>
  collect() |>
  filter(site_id == "CN09") |>
  ggplot(aes(x = sample_date, colour = monitoring_group)) +
  labs(title = "Enterococci results from water samples", y = "Enterococci (cfu per 100 mL)") +
  geom_segment(aes(xend = sample_date, y = 0, yend = numeric_value)) +
  geom_point(aes(y = numeric_value)) +
  geom_hline(yintercept = 240, colour = "white", size = 0.5, linetype = 2) +
  
  # Scales
  scale_colour_manual(values = c("routine" = city_green, "daily" = city_pink, "blue_flag" = city_blue)) +
  
  # Theme
  theme(
    legend.position = "bottom",
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(colour = "white"),
    legend.title = element_blank(),
    text = element_text(family = "Century Gothic"),
    plot.background = element_rect(fill = city_teal),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.grid = element_blank(),
    plot.title = element_text(colour = "white", hjust = 0.5, face = "bold"),
    legend.background = element_blank(),
    legend.text = element_text(colour = "white")
  )
```

```{r title-figure}
coast <- sites |>
  filter(site_id == "CN09") |>
  mutate(coast_label = str_c(coastline, " Coast", sep = "")) |>
  ggplot() +
  geom_text(aes(x = 0, y = 1, label = coast_label), angle = 90, size.unit = "mm", size = 10, family = "Century Gothic", colour = "white", fontface = "bold") +
  
  # Styles
  theme(
    plot.background = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_rect(city_blue)
  )
```

```{r site-name}
site_name <- sites |>
  filter(site_id == "CN09") |>
  mutate(coast_label = str_c(coastline, " Coast", sep = "")) |>
  ggplot() +
  geom_text(aes(x = 0, y = 1, label = site_description), angle = 90, size.unit = "mm", size = 10, family = "Century Gothic", colour = "white") +
  
  # Styles
  theme(
    plot.background = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_rect(fill = city_blue)
  )
```

```{r assemble-plot, fig.width = 6.4, fig.height = 9, dpi = 600}
sites_overview + theme(legend.position = "none") +
site_map +
site_name +
coast +
results_plot +
samples_plot +
hazen_plot +
exceedance_plot +
plot_layout(design = "
            CBBBB
            CBBBB
            CAAAA
            CAAAA
            DFFFF
            DEEEE
            ")




```

# Plot sample frequency in the reporting period
sites |>
  inner_join(results, by = "site_id") |>
  filter(between(sample_date, "2024-01-01", "2025-01-01"), monitoring_group %in% c("routine", "daily")) |>
  collect() |>
  filter(site_id == "CN11") |>
  group_by(month = floor_date(sample_date, "months"), site_id, monitoring_group) |>
  count() |>
  ggplot() +
  geom_col(aes(x = month, y = n, fill = monitoring_group), position = position_dodge(preserve = "single")) +
  scale_fill_discrete() +
  scale_x_date(breaks = months, date_labels = "%b")

ggplot() +
  
```